---
title: "ST309-week4"
output:
  pdf_document: default
  html_document: default
date: "2023-10-16"
---


```{r}

# install.packages('ISLR')
library('ISLR')
Colloge = College
attach(College)
```

(a)
```{r}
summary(College)
```
There are 565 private colleges, and 212 non-private colleges.

(b)

```{r}

```


```{r}
pairs(College[,1:10])
```

There are highly correlated pairs of variables such as (Apps, Accept), (Accept, Enroll), (Enroll,
P.Undergrad), (Top10perc, Top25perc). To make the plot less dense, we use the first 100 colleges
data to reproduce the plot pairs(College[1:100,1:10])


```{r}
pairs(College[1:100,1:10])
```

(c)
```{r}
plot(Private, Outstate, col=c("red","blue"), xlab="Private", ylab="Outstate
Tuition", main="Boxplots of out-of-state tuition of US colleges",
cex.main=2, cex.label=2, cex.axis=1.5)
```

(d)
```{r}
Elite=ifelse(Top10perc>50, "Yes","No")

Elite
```

```{r}
Elite=as.factor(Elite)
college=data.frame(College, Elite) # this new college has 19 variables
write.table(college, "college.txt")
```

```{r}
plot(Elite, Outstate, col=c("red","blue"), xlab="Elite", ylab="Outstate
Tuition", main="Boxplots of out-of-state tuition of US colleges",
cex.main=2, cex.label=2, cex.axis=1.5)
```

(e)
```{r}
par(mfrow=c(2,3))
hist(Apps, nclass=25)
hist(Accept, nclass=25)
hist(Enroll, nclass=25)
hist(Top10perc, nclass=25)
hist(F.Undergrad, nclass=25)
hist(Room.Board, nclass=25)

```
(f)(1)
```{r}
plot(S.F.Ratio, Grad.Rate) 
cor(S.F.Ratio, Grad.Rate) 
```

(f)(2)
```{r}
table(Elite, Private)
```
Note 65/(65 + 500) = 11.5% of private colleges are elite, while merely 13/(13 + 199) = 6.13% of
non-private colleges are elite

(f)(3)
```{r}
plot(Private, perc.alumni, xlab="Private", ylab="Percentage of alumni who donate", cex.label=2)
```

(f)(4)

There are different way to measure the importance. With the given info, the most relevant index
is the graduate rate. We could also include percentage of alumni who donate as an indication.

Since the faculty PhD rate is positively correlated with the graduation rate and the percentage
of donation, it is important.

```{r}
pairs(data.frame(PhD, Grad.Rate, perc.alumni))
cor(data.frame(PhD, Grad.Rate, perc.alumni))

```

5(b)

```{r}
JohnEvening=function(a, b) {
  X = runif(1, 0, 60)
  print(X)
  if((X>a)&(X<=b)) cat("visit mother", "\n")
  else if((X>(30+a))&(X<=(30+b))) cat("visit mother", "\n")
  else cat("visit girlfriend", "\n")
  
}

JohnEvening(5,15)
JohnEvening(5,15)

```

```{r}
ExpectationY = function(n, a, b){
  X = runif(n, 0, 60)
  Y = rep(0, n) # copy 0 n times in a vector
  for (i in 1:n){
    if( (X[i]>a) & (X[i]<=b) ){
      Y[i] = 1
      # cat("visit mother", "\n")
    }
    else if( (X[i]>(30+a)) & (X[i]<=(30+b)) ){
      Y[i] = 1
      # cat("visit mother", "\n")
    }
    else{
      Y[i] = 0
      # cat("visit girlfriend", "\n")
    }
  }
  return(30*mean(Y))
}

set.seed(62537)
ExpectationY(10000,5,15)
```


```{r}
# set.seed(65999)
ExpectationY(100000,5,15)
```


Why use sort?
Without sort, need three conditions:
X_1 + X_2 > X_3
X_2 + X_3 > X_1
X_1 + X_3 > X_2

after sort, just need one:
X_1 < X_2 < X_3

```{r}
prob3angle = function(n) {
  X = matrix(runif(3*n), ncol=3)
  for(i in 1:n) X[i,]=sort(X[i,]) # put each row in ascending order
  Y = X[,1][(X[,1]+X[,2])>X[,3]] # remove the row that three values cannot be lengths for 3 sides.
  cat("The probability is:",length(Y)/n,"\n")
}

set.seed(1763)
prob3angle(10000)
set.seed(1763)
prob3angle(10000)

prob3angle(10000)
prob3angle(10000)
prob3angle(10000)
prob3angle(10000)
prob3angle(10000)


```
```{r}
prob3angle(4)
```

use loops
```{r}
Prob3angles = function(n){
  X = matrix(runif(3*n), ncol=3)
  Y = rep(0, n)
  for(i in 1:n) X[i,] = sort(X[i,])
  for(i in 1:n){
    if( X[i,1]+X[i,2]>X[i,3] ){
      Y[i] = 1
    }
  }
  return(mean(Y))
}

Prob3angles(10000)
Prob3angles(10000)
Prob3angles(10000)
Prob3angles(10000)

```


```{r}
set.seed(10000)
rnorm(5)

set.seed(9999)
rnorm(5)
```

