---
title: "ST309 Week 3"
output:
  pdf_document: default
  html_document: default
date: '2023-10-09'
---

## 3.2 Functions

# Elements in defining a function: input, output, how to get output from input.

# One can set default input.


```{r}
Addfunction = function(a,b){
  return(a+b)
}

# Addfunction(2)
Addfunction(2,4)
# Addfunction()
```

# Side effect: sometimes we do not need any output, instead, we want the function to draw a plot for example.

```{r}
Meanfunction = function(a){
  plot(a) 
  return(sum(a)/length(a))
}

a = 5:20
Meanfunction(a)

```

# *recursion (advanced topic)

```{r}
Fab = function(n){
  if (n==1 || n==2){
    return(1)} 
  else return(Fab(n-1) + Fab(n-2))
}
```


```{r}
Fab(20)
```




## 3.3 Loops

# For loop: when you know exactly the number of iteration steps

```{r}

# a = c(1,2,3,2,4,5)
a = 1:100

# sum(a)

summation = 0
l = length(a)
for (i in 1:l){
  summation = summation + a[i]
  # print(i)
}

a = 1:100
i = 1
while (i<l){
  summation = summation + a[i]
  i = i + 1
}

summation
```


```{r}
a = 1:100
l = length(a)
i = 1
summation = 0
while (i<=l){
  summation = summation + a[i]
  i = i + 1
}

summation
```

# While: when you know the end condition but not the number of iteration steps

```{r}
j <- 1
while (j<6){
  print(j)
  j = j+1
}
```

# repeat: using 'break' to specify terminating condition

(I never used 'repeat'.)

```{r}
j = 1
repeat{
  print(j)
  j <- j+1
  if (j>5) break
}
```

## 3.4 If else

```{r}
a = 1:100
l = length(a)
summation = 0
for (i in 1:l){
  if (i%2 == 1){
    summation = summation + a[i]
  }
}

```


```{r}
x = 11
# ans = 0

if (x>=10){
  ans = x + 10}


ans
```

```{r}
x = 0

if (x>=10){ans = x+10} else {ans=x}

ans
```

# change line: Curly bracket needed

```{r}
x = 0
{if (x>=10) ans = x+10 
  else ans=x}
ans
```

# else if: More than 2 situations

```{r}
x = 1
{if (x>=10) ans = x+10
  else if (x>0) {ans = x+1
      ans = ans + 1 }
  else if (x>-1) ans = x
  else if (x>-1.5) ans = ans
  else ans = x - 1}
ans
```

# ifelse(condition, vec1, vec2) can be applied to a vector

```{r}
setwd('/Users/wyt233/Library/CloudStorage/OneDrive-LondonSchoolofEconomics/ST309 2024/week3')
mk2nd = read.table('marks2.txt', header=T, row.names = 1)
attach(mk2nd)
pf <- ifelse(exam1 + exam2 + exam3>=60, 'PASS', 'FAILED')
pf

```

## 3.5 Avoiding loops

```{r}
# exam1+exam2+exam3
pf2 = apply(mk2nd, 1, sum)

# 1 means applying to all rows
# 2 means applying to all columns

pf2
```

```{r}
mat = matrix(1:24, 8, 3)
mat
# apply(mat, 2, mean)

#apply(mat, 1, mean)
# about the second argument
# 1 - for row
# 2 - for column
# apply(mat, 1, sum)
# apply(mat, 2, max)
# customise the function
# apply(mat, 2, function(x) x[1]+x[2] )
```

# Q1

```{r}
test = function(x){
  p = length(x)
  
  res = matrix(0,p,p) # initialisation
  
  for (i in 1:p){
    for (j in 1:p){
      res[i,j] = x[i] - x[j]
    }
  }
  return(res)
}

test(1:10)
```

# An example in numerical analysis

A typical situation to use 'while' rather than 'for':

```{r}
left = -1
right = 0
step = 0


while ( right - left > 0.000000001){
  new = (left + right)/2
  f = exp(new) + new
  if (f > 0){
    right = new
  }
  else if (f < 0){
    left = new
  }
  step = step + 1
}
```
